---
title: "reproResearch"
author: "suszanna"
date: "10/19/2020"
output: html_document
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
#install.packages("ggplot2")
```

## Reproducible Research: A Quantified Analysis of Activity Monitoring Data

Collected from an activity monitoring device, the 'activity.csv' dataset contains 17568 observations and 3 variables: steps, time and interval. The dataset consists of two months of data from a single individual and samples the number of steps taken in 5 minute intervals.  The data contain imputable missing values.  Computed and plotted are (1) a histogram of the total steps taken each day, (2) a time series plot of the average daily number of steps taken, (3) a histogram of the total number of steps taken each day after missing values are imputed, and (4) a panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends. This report contains the 8 points of interest below, as well as all of the R code needed to reproduce the results (numbers, plots, etc.).

When you click the 'Knit to HTML' button, a document will be generated that includes both content and the output of any embedded R code chunks within the document. The steps to prepare the data and to reproduce the 4 plots follow.

1) Read in the dataset, view the data
2) plot1: Show Histogram of the Total Number of Steps taken each Day
3) Calculate Mean and Median Number of Steps taken each Day
4) plot 2: Show a Time Series Plot of the Average Number of Steps Taken
5) Identify the 5-minute Interval that, on Average, contains the Maximum Number of steps
6) Show the Code that Describes a Strategy for Imputing Missing Data
7) plot 3: Show a Histogram of the Total Number of Steps taken each day after Missing Values are Imputed
8) plot 4: Show a Panel Plot comparing the Average Number of Steps taken per 5-minute interval Across Weekdays and Weekends

9) Show all of the R code needed to reproduce the results (numbers, plots, etc.) is included in this report.

```{r activity, echo=TRUE}

#1) Read in the dataset, view the data
activity<-read.csv(unzip("activity.zip"))
dim(activity)
str(activity)
head(activity)
tail(activity)
summary(activity)

#Two ways to count original number of missing values
sum(is.na(activity))
dim(activity[activity$steps == "NA", ])[1]

par(mfrow = c(1, 1))

#2) plot1: Show Histogram of the total number of steps taken each day
# 'total_step' - total number of steps taken per day -ignore NAs
totalsteps <- aggregate(steps ~ date, data = activity, sum, na.rm = TRUE)
head(totalsteps)

hist(totalsteps$steps, breaks = 20, 
     main = "1 Total Steps Taken Each Day",
     col = "blue", border = "grey", xlab = "Steps", ylab = "Frequency", axes = FALSE)
axis(1)
axis(2, las = 1)

#3) mean and median total steps taken each day
#summary(total_step)
summary(totalsteps$steps)
mean(totalsteps$steps)
median(totalsteps$steps)


#4) Plot 2: Show time series plot of the average number of steps taken
avgsteps <- aggregate(steps ~ interval, data = activity, mean, na.rm = TRUE)
plot(avgsteps$interval, avgsteps$steps, type = "l", lwd = 2, col = "blue",
     main = "2 Time Series: Avg Number of Steps Taken", axes = FALSE,
     xlab = "5-minute interval", ylab = "Avg number of steps")
axis(1)
axis(2, las = 1)

#5) Identify the 5-minute interval that, on average, contains the maximum number of steps
# max num steps = 206+
#which.max(avg_step$steps)
#avg_step$steps[which.max(avg_step$steps)]

# this is the interval number we're looking for
avgsteps$interval[which.max(avgsteps$steps)]


#6) Show code to describe a strategy to impute (replace) missing data
imputed <- activity # create dataset to impute ('imputed')
for (i in avgsteps$interval) {
    imputed[imputed$interval == i & is.na(imputed$steps), ]$steps <- 
        avgsteps$steps[avgsteps$interval == i]
}

# no NAs- all are now imputed (replaced)
head(imputed) 
tail(imputed)
# Count of missing values in the imputed dataset
sum(is.na(imputed))
dim(imputed[imputed$steps == "NA", ])[1]

#7) Plot 3: Show histogram of the total number of steps taken each day after missing values are imputed
#totalsteps_imputed <- aggregate(steps ~ date, data = imputed, sum, na.rm = FALSE)
totalsteps_imputed <- aggregate(steps ~ date, data = imputed, FUN=sum)
hist(totalsteps_imputed$steps, breaks = 20, 
     main = "3 Total Steps Taken Each Day -imputed",
     col = "blue", border = "red", xlab = "Steps", ylab = "Frequency", axes = FALSE)
axis(1)
axis(2, las = 1)

#8) Plot 4: Show panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```

```{r imputed$dateType, echo=TRUE}

library(lubridate)
imputed$dateType <-  ifelse(as.POSIXlt(imputed$date)$wday %in% c(0,6), 'weekends', 'weekdays')

library(ggplot2)
meanimputed <- aggregate(steps ~ interval + dateType, data=imputed, FUN=mean)
ggplot(meanimputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avg number of steps")+
    ggtitle("4 Avg Steps Taken- weekdays v. weekends")+
    theme(plot.title = element_text(size=15, face="bold", hjust = 0.5))
    #geom_point(size=1) +

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
